### 2024-07-01 test/main.tex
上传

\documentclass{beamer}
\usetheme{Madrid}
\usefonttheme{professionalfonts}
\AtBeginSection[]{
\begin{frame} 
    \frametitle{Contents}
    \tableofcontents[currentsection]
\end{frame}
}
\title[CHY formula]{Scattering of massless particles: scalars, gluons and gravitons}
\subtitle{Freddy Cachazo, Song He, Ellis Ye Yuan}
\author{Su Yingze}
\institute{Nagoya University}
\date[7 23rd 2024]{July 23rd 2024} 
\begin{document}
\frame{\titlepage}

\section{Why the S-matrix is so important?}
\begin{frame}{Locality and Unitarity}
    aa
\end{frame}
\section{Review of some pertinent results}
\begin{frame}{Gauge theory color structure}
    At tree level, with particles in the adjoint representaion of gauge group SU(N), the amplitude can be 
decomposed as
\textcolor{red}{\begin{equation*}
    \mathcal{A}_n^{tree}(1,2,3,\dots,n)=\sum _{\mathcal{P}(2,3,\dots,n)}Tr[T^{a_1}T^{a_2}T^{a_3}\dots
    T^{a_n}]A_n^{tree}[1,2,3,\dots,n]
\end{equation*}}

here we ommit the coupling constant g, and $A_n^{tree}$ is called tree-level color-ordered partial amplitude.
Notice that this basis includes \textcolor{red}{$(n-1)!$} independent amplitudes.
\end{frame}
\begin{frame}
    Color-ordered partial amplitudes satisfy a set of well-known properties,
    \begin{itemize}
        \item Cyclic:\quad $$A_n^{tree}[1,2,3,\dots,n]=A_n^{tree}[2,3,\dots,n,1]$$
        \item Reflection: \quad $$A_n^{tree}[1,2,3,\dots,n]=(-1)^nA_n^{tree}[n,\dots,3,2,1]$$
        \item "photon" decoupling: \quad $$\sum_{\sigma\in \text{cyclic}}A_n^{tree}[1,\sigma(2,3,\dots,n)]=0$$
        \item KK(Kleiss-Kuijf) relation:\quad $$A_n^{tree}[1,{\alpha},n,{\beta}]=(-1)^{n_\beta}
        \sum_{\left\{\sigma\right\}_i\in OP(\left\{\alpha\right\},\left\{\beta\right\}^T)}A_n^{tree}[1,\left\{\sigma\right\}_i,n]$$
    \end{itemize}
    where the OP means ordered permutaions, that is all permutations of$\left\{\alpha\right\},\left\{\beta\right\}^T$
    that matains the order of individual elements of each set.
\end{frame}
\begin{frame}
    For example, a five point amplitude $A_n^{tree}(1,\{2,3\},5,\{4\})$, we have
\begin{equation*}
    A_n^{tree}[1,2,3,5,4]=-A_n^{tree}[1,2,3,4,5]-A_n^{tree}[1,2,4,3,5]-A_n^{tree}[1,4,2,3,5]
\end{equation*}
The other five point relations can be obtained by permuting legs 2,3,4 and using cyclic and reflection properties.
\\ \hspace*{\fill}\\
This means that the six amplitudes $A_n^{tree}(1,\mathcal{P}\{2,3,4 \},5 )$ form a basis of remaining five-point partial amplitudes.
\textcolor{red}{More generally, for multiplicity n, the KK relation can be used to rewrite any color-ordered partial amplitude in terms of only $(n-2)!$
basis partial amplitudes, where two legs are fixed(usually choose 1 and n).}
\end{frame}
\begin{frame}{A Similar Structure}
    It can be proved that tree level amplitudes can also be decomposed like 
\begin{equation*}
    \mathcal{A}_n^\text{tree}=\sum_{\sigma\in S_{n-2}}\tilde{f}^{a_1a_{\sigma_1}b_1}\tilde{f}^{b_1a_{\sigma_2}b_2}\cdots\tilde{f}^{b_{n-3}a_{\sigma_{n-2}}a_n}\tilde{A}_n(1,\sigma_1,\sigma_2,\ldots,\sigma_{n-2},n)
\end{equation*}
It is easy to realize that the number of basis amplitudes is also \textcolor{red}{$(n-2)!$}.
\end{frame}
\begin{frame}
    \begin{Large}
        \textcolor{red}{How can we realize the existence of this basis?}
    \end{Large}
\\ \hspace*{\fill}\\
We just need to notice the contribution from ladder diagram
\begin{figure}[htb]
    \centering
    \includegraphics[width=1\linewidth]{1.png}
\end{figure}
And any (three vertex) diagram, by using the Jacobbi identity like 
\begin{figure}[htb]
    \centering
    \includegraphics[width=1\linewidth]{2.png}
\end{figure}
can be transformed to the ladder diagram. It has been proved that $\textcolor{red}{\tilde{A}_n(\dots)}$ here is identical to $\textcolor{red}{A_n^{tree}[\dots]}$ in the trace basis.
\end{frame}
\begin{frame}{Color-Kinematics Duality and BCJ Relation}
The Color-Kinematics Duality here is not the complete version, but we can sketch some point from lower point case
\end{frame}

\section{CHY Formula}
\begin{frame}{Scattering equations}
    It has benn proposed that there is connection between the scattering data of n massless particles and the n-punctured sphere from a rational map
    \begin{equation*}
        k_\mu^a=\frac{1}{2\pi i}\oint_{|z-\sigma_a|}dz\frac{p^\mu(z)}{\prod ^n_{b=1}(z-\sigma_b)}
    \end{equation*}

    To discribe the n-punctured sphere more properly, we can introduce the Riemann sphere as
    \begin{equation*}
        \mathbb{CP}^1\cong \mathcal{S}^2\cong \mathbb{C}\cup\{\infty\}
    \end{equation*}
    and n-punctured Riemann sphere can be discribed by $SL(2,\mathbb{C})$ affine coordinates ${\sigma_1,\sigma_2,\ldots,\sigma_n}$, that is to say we have a equvilance relation
    \begin{align*}
        \{\sigma_1,\sigma_2,\ldots,\sigma_n\}\sim\{\psi(\sigma_1),\psi(\sigma_2),\ldots,\psi(\sigma_n)\},\\\psi(\sigma):=\frac{\alpha\sigma+\beta}{\gamma\sigma+\delta},\quad\alpha,\beta,\gamma,\delta\in\mathbb{C},\alpha\delta-\beta\gamma=1
    \end{align*}
    because of the redundancy of 
\end{frame}
\begin{frame}
    From this map, we can easily obtain the main ingredients of this report \\

    \begin{block}{\textcolor{red}{Scattering equations}}
        \begin{equation*}
            \sum_{b\neq a } \frac{s_{ab}}{\sigma_a-\sigma_{b}}=0,\qquad a\in \{1,\dots,n\}
        \end{equation*}
    \end{block}
    It has been proved that the number of solutions in any dimension is $\alert{(n-3)!}$, and only \alert{n-3} of the equtions are independent, so we can rewrite the scattering equations as following 
    \begin{equation*}
        \sum_{b\neq a } \frac{s_{ab}}{\sigma_a-\sigma_{b}}=0,\quad a\in \{4,5,\dots,n\}\quad and\quad  \sigma_1\rightarrow\infty,\sigma_2=0,\sigma_3=1
    \end{equation*}
\end{frame}
\begin{frame}{KLT Relation}
    So-called KLT relation, in the language of field theory, refers to the decomposition of \alert{gravity amplitudes} to \alert{two gauge theory amplitues}. 
    \begin{table}[]
        \resizebox{\textwidth}{!}{
        \begin{tabular}{|l|l|l|l|l|}
        \hline
         \(\mathcal{N}\) & Factors & Supergravity   \\ \hline
         8 &   \(\mathcal{N}=4SYM\otimes \mathcal{N}=4SYM\)      &   pure \(\mathcal{N}=8 \)SG             \\ \hline
         6 &   \(\mathcal{N}=4SYM\otimes \mathcal{N}=2SYM\)      &   pure \(\mathcal{N}=6 \)SG             \\ \hline
         5 &   \(\mathcal{N}=4SYM\otimes \mathcal{N}=1SYM\)      &   pure \(\mathcal{N}=5 \)SG             \\ \hline
         4 &   \(\mathcal{N}=4SYM\otimes (\mathcal{N}=0YM+n_\nu scalars)\)      &    \(\mathcal{N}=4 \)SG,\(n_\nu\)vector multiplets             \\ \hline
         4 &   \(\mathcal{N}=2SYM\otimes \mathcal{N}=2SYM\)      &    \(\mathcal{N}=4 \)SG,2 vector multiplets            \\ \hline
         3 &   \(\mathcal{N}=2SYM\otimes \mathcal{N}=1SYM\)      &    \(\mathcal{N}=3 \)SG,1 vector multiplet           \\ \hline
         2 &   \(\mathcal{N}=2SYM\otimes (\mathcal{N}=0YM+n_\nu scalars)\)      &    \(\mathcal{N}=2 \)SG,\(n_\nu\) multiplets+1 vector multiplets            \\ \hline
         2 &   \(\mathcal{N}=1SYM\otimes \mathcal{N}=1SYM\)       &    \(\mathcal{N}=2 \)SG ,1 hypermultiplet               \\ \hline
         1 &   \(\mathcal{N}=1SYM\otimes (\mathcal{N}=0YM+n_\nu scalars)\)      &     \(\mathcal{N}=1 \)SG,  \(n_\nu\) vector and 1 chiral multiplet           \\ \hline
        \end{tabular}
        }
        \end{table}
\end{frame}
\begin{frame}{KLT Orthogonality}
    $\bullet$ KLT orthogonality is a striking property of the solutions to scattering equations.
    \begin{block}{Proposition 1}
        \begin{equation*}
            \frac{(i,j)}{(i,i)^{\frac12}(j,j)^{\frac12}}=\delta_{ij}
        \end{equation*}
    \end{block}

\end{frame}
\begin{frame}
    First we need to define the Jacobian matrix associated to the scattering equations

        \begin{equation*}
            \boxed{
            \label{jacobbi}
            \Phi_{ab}\equiv\partial\left(\sum_{c\neq a}\frac{s_{ac}}{\sigma_a-\sigma_c}\right)/\left(\partial\sigma_b\right)=\begin{cases}\frac{s_{ab}}{(\sigma_a-\sigma_b)^2},&a\neq b,\\-\sum_{c\neq a}\Phi_{ac}, &a=b\end{cases}
            }
        \end{equation*}
    As mentioned above only n-3 of the scattering equations are independent so the matrix $\Phi$ has \alert{rank n-3}.  (This matrix was first encountered in the gravity amplitudes constructed from gauge theory using KLT relation ) 
\end{frame}
\begin{frame}
    Consider a generalization of $\Phi_{ab}$ 
    \begin{equation*}
        \Psi_{ab,a\neq b}\equiv\frac{s_{ab}}{(\sigma_a-\sigma_b)(\sigma_a^{\prime}-\sigma_b^{\prime})},\quad\Psi_{aa}\equiv-\sum_{c\neq a}\Psi_{ac}.
    \end{equation*}
    \begin{block}{Proposition 2}
        \begin{equation*}
            \operatorname{rank}\Psi(\{\sigma\},\{\sigma^{\prime}\})=\begin{cases}n-4, \{\sigma\}\neq\{\sigma^{\prime}\}\\n-3, \{\sigma\}=\{\sigma^{\prime}\}\end{cases}
        \end{equation*}
        $\sigma$ and $\sigma^\prime$ are assmused to be solutions to scattering equation.
    \end{block}
\end{frame}
\begin{frame}{Prove of KLT orthogonality}
    For the purpose of proving KLT orthogonality, we can construct a n! dimension vector for each solution
    \begin{equation*}
        \frac1{(\sigma_{\omega(1)}-\sigma_{\omega(2)})(\sigma_{\omega(2)}-\sigma_{\omega(3)})\cdots(\sigma_{\omega(n)}-\sigma_{\omega(1)})}
    \end{equation*}
    Not so obvious is the fact that we can fix the positon of 3 labels, which we choose 1,n-1,n, give rise to the KK relation and BCJ relation.
    \\ \hspace*{\fill}\\
    Now the vectors become (n-3)! dimension, and even aftering selecting three lables, we still have the freedom of where to put them. Here we only use two choices :
    \begin{equation*}
        \textcolor{red}{(1,\omega(2),\dots,\omega(n-2),n-1,n)\quad and \quad (1,\omega(2),\dots,\omega(n-2),n,n-1 )}
    \end{equation*}
\end{frame}
\begin{frame}
    The corresponding two vectors are 
    \begin{align*}V(\omega)&=\frac1{(\sigma_1-\sigma_{\omega(2)})\cdots(\sigma_{\omega(n-2)}-\sigma_{n-1})(\sigma_{n-1}-\sigma_n)(\sigma_n-\sigma_1)},\\U(\omega)&=\frac1{(\sigma_1-\sigma_{\omega(2)})\cdots(\sigma_{\omega(n-2)}-\sigma_n)(\sigma_n-\sigma_{n-1})(\sigma_{n-1}-\sigma_1)}.
    \end{align*}
    In this language, we can construct a bilinear form 
    \begin{equation*}
        S[\alpha|\beta]=\prod^{n-2}_{i=2}\left(s_{1,\alpha_i}+\sum_{j=2}^{i-1}
        \theta(\alpha(j),\alpha(i))_\beta s_{\alpha(j),\alpha(i)}\right)
    \end{equation*}
    where $\alpha,\beta\in S_{n-3}$,$\theta(i,j)_\beta=1$if the order of i,j is the same in both permutations 
    $\alpha(2,3,\dots,n-2)$ and $\beta(2,3,\dots,n-2)$, and 0 otherwise.
    S is usually called \alert{Momentum Kernel}.
\end{frame}
\begin{frame}
    Given any two solutions of scatering equations,
    \begin{equation*}
        \{\sigma_1^{(i)},\sigma_2^{(i)},\dots,\sigma_n^{(i)}\}\quad and \quad \{\sigma_1^{(j)},\sigma_2^{(j)},\dots,\sigma_n^{(j)}\}
    \end{equation*}
    define two vectors, $V(\alpha)^{(i)}$ and $U(\beta)^{(j)}$, i,j are choices of solutions and $\alpha,\beta$ are the choices of permutaions,
    the number of both is \alert{(n-3)!}.
    \\ \hspace*{\fill}\\
    A natural inner product can be defined as
    \begin{equation*}
        (i,j):=\sum_{\alpha,\beta\in S_{n-3}}V^{(i)}(\alpha)S[\alpha|\beta]U^{(j)}(\beta)
    \end{equation*}
    Knowing all definetions above, we can proceed to prove KLT orthogonality.
\end{frame}
\begin{frame}
    The starting point is to notice that
    \begin{equation*}
        \frac{(i,j)}{(i,i)^{\frac12}(j,j)^{\frac12}}=\delta_{ij}
    \end{equation*}
    is clearly invariant under $SL(2,\mathbb{C})\times SL(2,\mathbb{C})$. Partially fixing both $SL(2,\mathbb{C})$
    redundancies with convenient choice $\sigma_{n-1}^{(i)}=\sigma_n^{(j)}=\infty$ and $\sigma_n^{(i)}=\sigma_{n-1}^{(j)}=1$
    and define 
    \begin{align*}
            K_n(\{\sigma\},\{\sigma'\})\equiv\sum_{\alpha,\beta\in S_{n-3}}&\frac1{\sigma_{1,\alpha(2)}\ldots\sigma_{\alpha(n-3),\alpha(n-2)}}S[\alpha|\beta]\\
            &\frac1{\sigma_{1,\beta(2)}^{\prime}\ldots\sigma_{\beta(n-3),\beta(n-2)}^{\prime}}
    \end{align*}
    The motivation for this definition is that $K_n$ appears in the numerator of KLT orthogonality.
\end{frame}
\begin{frame}
    It is also convenient to define an auxiliary co-rank one $(n-2)\times(n-2)$ matrix $\psi^{(n)}$
    \begin{equation*}
        \psi_{ab,a\neq b}=\frac{s_{ab}}{\sigma_{ab}\sigma_{ab}^{\prime}},\quad\psi_{aa}=-\sum_{b\neq a}\psi_{ab},\quad a,b=1,\ldots,n{-}2
    \end{equation*}
    It can be proven that any $(n-3)\times(n-3)$ minors of $\psi^{(n)}$ are the same, and we denote such a minor as
    ${\det}^\prime\psi^{(n)}$ , that is to say, the determinat of the matrix after removing any row and colum.
    \begin{block}{Proposition 3}
        The two functions defined above are identical up to a sign.
        \begin{equation*}
            K_n(\{\sigma\},\{\sigma'\})=(-1)^n{\det}^\prime\psi^{(n)}
        \end{equation*}
    \end{block}
\end{frame}
\begin{frame}
    The final step is put all pieces together. With the choice  $\sigma_{n-1}^{(i)}=\sigma_n^{(j)}=\infty$ and $\sigma_n^{(i)}=\sigma_{n-1}^{(j)}=1$,
    we have 
    \begin{equation*}
        \frac{(i,j)}{(i,i)^{\frac12}(j,j)^{\frac12}}=\frac{K_n(\{\sigma^{(i)}\},\{\sigma^{(j)}\})}{K_n^{\frac12}(\{\sigma^{(i)}\},\{\sigma^{(i)}\})K_n^{\frac12}(\{\sigma^{(j)}\},\{\sigma^{(j)}\})}
    \end{equation*}
    In addition, one finds that the minor of $\psi$ obtained by removing the first row and colum is identical to 
    that of $\Psi(\{\sigma\},\{\sigma^{\prime}\})$ after removing rows and colums $\{1,n-1,n\}$. We denote them respectively
    $|\psi^{(n)}|^1_1$ and $|\Psi|^{1,n-1,n}_{1,n-1,n}$. Then,
    \begin{align*}
        \frac{(i,j)}{(i,i)^{\frac12}(j,j)^{\frac12}}&=\frac{K_n(\{\sigma^{(i)}\},\{\sigma^{(j)}\})}{K_n^{\frac12}(\{\sigma^{(i)}\},\{\sigma^{(i)}\})K_n^{\frac12}(\{\sigma^{(j)}\},\{\sigma^{(j)}\})}\\
        &=\frac{(-1)^n|\psi^{(n)}|^1_1}{{(-1)^n|\psi^{(n)}|^1_1}^{\frac{1}{2}}{|\psi^{(n)}|^1_1}^{\frac{1}{2}}}\\
        &=\frac{|\Psi(\{\sigma^{(i)}\},\{\sigma^{(j)}\})|_{1,n-1,n}^{1,n-1,n}}{(|\Psi(\{\sigma^{(i)},\sigma^{(i)}\}|_{1,n-1,n}^{1,n-1,n})^{\frac12}(|\Psi(\{\sigma^{(j)}\},\{\sigma^{(j)}\}|_{1,n-1,n}^{1,n-1,n})^{\frac12}}
    \end{align*}
\end{frame}
\begin{frame}
    Fianlly, we just need to use Proposition 2. 
    \begin{itemize}
        \item If $i=j$, the rank of matrix $\Psi$ is $n-3$ and the minor is nonzero, we obtain
        \begin{equation*}
            \frac{(i,i)}{(i,i)^{\frac12}(i,i)^{\frac12}}=\frac{|\Psi(\{\sigma^{(i)}\},\{\sigma^{(j)}\})|_{1,n-1,n}^{1,n-1,n}}{|\Psi(\{\sigma^{(i)}\},\{\sigma^{(j)}\})|_{1,n-1,n}^{1,n-1,n}}=1
        \end{equation*}
        \item If $i\neq j$, the rank of matrix is $n-4$, so any minor with volume more than $n-4$ equals 0.
        \begin{equation*}
            |\Psi(\{\sigma^{(i)}\},\{\sigma^{(j)}\})|_{1,n-1,n}^{1,n-1,n}=0\Rightarrow \frac{(i,j)}{(i,i)^{\frac12}(j,j)^{\frac12}}=0
        \end{equation*}
    \end{itemize}
    Up to now, we conclude the proof of KLT orthogonality.
\end{frame}
\begin{frame}
    \frametitle{Attempt to construct S-matrix --- Towards CHY}
    Thanks to the excellent properties of scattering equations, it is very tempting to propose that the solutions
    to scattering equations should be used to construct scattering amplitudes.
    \\ \hspace*{\fill}\\
    The first two constructed are YM and gravity amplitudes in any dimensions
    \begin{align*}
        M_n^{\mathrm{YM}}(1,2,\dots,n)&=\int\frac{d^n\sigma}{\text{vol SL}(2,\mathbb{C})}\prod_a{'}\delta\bigg(\sum_{b\neq a}\frac{s_{ab}}{\sigma_{ab}}\bigg)\frac{E_n(\{k,\epsilon,\sigma\})}{\sigma_{12}\dots\sigma_{n1}},\\
        M_n^{\mathrm{gravity}}&=\int\frac{d^n\sigma}{\text{vol SL}(2,\mathbb{C})}\prod_a{'}\delta\bigg(\sum_{b\neq a}\frac{s_{ab}}{\sigma_{ab}}\bigg)E_n(\{k,\epsilon,\sigma\})^2
    \end{align*}
    The measure is defined as following
    \begin{equation*}
        \prod_{a}{}^{\prime}\delta{\left(\sum_{b\neq a}\frac{s_{ab}}{\sigma_{ab}}\right)}{:}=\sigma_{ij}\sigma_{jk}\sigma_{ki}\prod_{a\neq i,j,k}\delta{\left(\sum_{b\neq a}\frac{s_{ab}}{\sigma_{ab}}\right)}
    \end{equation*}
\end{frame}
\begin{frame}
    The reason we extract 3 indices from delta equation is the fact that only $n-3$ scattering equations are independent.
    This from can be proved to be \alert{independent of choice of $i,j,k$}, therefore permutaion invariant.
    We also have
    \begin{equation*}
        \sigma_a\to\frac{\alpha\sigma_a+\beta}{\gamma\sigma_a+\delta}:\quad d\mu_n\to\prod_{a=1}^n(\gamma\sigma_a+\delta)^{-4}d\mu_n
    \end{equation*}
    $E_n(\{k,\epsilon,\sigma\})$ itself is permutaion invariant with resprct to $\sigma_a,k_a^\mu$ and $\epsilon_a^\mu$. The $SL(2,\mathbb{C})$ invariance of amplitude also 
    constraints the form of $E_n(\{k,\epsilon,\sigma\})$
    \begin{equation*}
        \sigma_a\to\frac{\alpha\sigma_a+\beta}{\gamma\sigma_a+\delta}:\quad E_n(\{k,\epsilon,\sigma\})\to E_n(\{k,\epsilon,\sigma\})\prod_{a=1}^n(\gamma\sigma_a+\delta)^2
    \end{equation*}
\end{frame}
\begin{frame}{The form of measure}
    It is worth to computed the measure explictly.Aftering "gauge fixing" the $SL(2,\mathbb{C})$ redundancy, one finds 
    \begin{equation*}
        \int\prod_{c\neq p,q,r}d\sigma_c(\sigma_{pq}\sigma_{qr}\sigma_{rp})(\sigma_{ij}\sigma_{jk}\sigma_{ki})\prod_{a\neq i,j,k}\delta\biggl(\sum_{b\neq a}\frac{s_{ab}}{\sigma_{ab}}\biggr)
    \end{equation*}
    The delta functions completely localize all integrals and the answer is evaluating a Jacobian defined above.
    \begin{equation*}
            \Phi_{ab}\equiv\partial\left(\sum_{c\neq a}\frac{s_{ac}}{\sigma_a-\sigma_c}\right)/\left(\partial\sigma_b\right)=\begin{cases}\frac{s_{ab}}{(\sigma_a-\sigma_b)^2},&a\neq b,\\-\sum_{c\neq a}\Phi_{ac}, &a=b\end{cases}
    \end{equation*}
    Then, we obtain the measure
    \begin{equation*}
            \boxed{
            \sum_{\{\sigma\}\in\mathrm{solutions}}\frac{(\sigma_{pq}\sigma_{qr}\sigma_{rp})(\sigma_{ij}\sigma_{jk}\sigma_{ki})}{|\Phi|_{pqr}^{ijk}}}
    \end{equation*}
\end{frame}
\begin{frame}
    Always denoted by 
    \begin{equation*}
        \boxed{
        \det{'}\Phi:=\frac{|\Phi|_{pqr}^{ijk}}{(\sigma_{pq}\sigma_{qr}\sigma_{rp})(\sigma_{ij}\sigma_{jk}\sigma_{ki})}}
    \end{equation*}
    $|\Phi|_{pqr}^{ijk}$ means that we need to delete the rows $\{i,j,k\}$ and the colums $\{p,q,r\}$, of course it is free to choose which inedex refers
    to row or colum ($\Phi$ is a symmetric matrix).
\end{frame}
\begin{frame}
    \frametitle{The form of $E_n(\{k,\epsilon,\sigma\})$}
    In order to present the explict form of $E_n(\{k,\epsilon,\sigma\})$, first define the following $2n\times2n$ antisymmetric matrix 
    \begin{equation*}
        \Psi=\begin{pmatrix}
            A & -C^T \\
            C & B 
        \end{pmatrix}
    \end{equation*}
    where $A,B$ and $C$ are $n\times n$ matrices, defined as 
    \begin{equation*}
        A_{ab}=\begin{cases}\frac{s_{ab}}{\sigma_a-\sigma_b}&a\neq b,\\0&a=b,\end{cases} \quad 
        B_{ab}=\begin{cases}\frac{\epsilon_a \cdot \epsilon_b}{\sigma_a-\sigma_b}&a\neq b,\\0&a=b, \end{cases} \quad
    \end{equation*}
    \begin{equation*}
        C_{ab}=\begin{cases}\frac{\epsilon_a\cdot k_b}{\sigma_a-\sigma_b}&a\neq b,\\-\sum_{c\neq a}\frac{\epsilon_a\cdot k_c}{\sigma_a-\sigma_c}&a\neq b.\end{cases}
    \end{equation*}
\end{frame}
\begin{frame}
    The first important observation is that while the Pfaffian of $\Psi$ is 0, but after removing any rows $i,j$ and colums $i,j$
    with $1\leq i< j\leq n$, the new matrix $\Psi_{ij}^{ij}$ have nonzero Pfaffian and we define the corresponding reduced Pfaffian as 
    \begin{equation*}
            \boxed{
            \mathrm{Pf}^{\prime}\Psi:=\frac{(-1)^{i+j}}{(\sigma_i-\sigma_j)}\mathrm{Pf}(\Psi_{ij}^{ij})}
    \end{equation*}
    It can be proved that the reduced Pfaffian is invariant under permutaion of \alert{particle labels}.

    \begin{alertblock}{Pfaffian}
        Pfaffian is defined for antisymmetric matrix, usually in two ways as following
        \begin{itemize}
            \item \begin{equation*}
                \mathrm{Pf}(A)^2=\det A
            \end{equation*}
            \item \begin{equation*}
                \mathrm{Pf}(A)=\frac{1}{2^n n!}\sum_{\sigma \in S_{2n}}sgn(\sigma)\prod_{i=1}^n a_{\sigma_{(2i-1)},\sigma(2i)}
            \end{equation*}
        \end{itemize}
    \end{alertblock}
\end{frame}
\begin{frame}
    Write down the proposal
    \alert{
    \begin{equation*}
        E_n(\{k,\epsilon,\sigma\})=\mathrm{Pf}^\prime \Psi(k,\epsilon,\sigma)
    \end{equation*}}
\end{frame}

\end{document}