### 2024-09-13 Paper Category/main.tex
New

\documentclass[a4paper,12pt]{article}
\usepackage[margin=1in]{geometry}
\title{}
\author{}
\begin{document}
    \begin{itemize}
        \item %\cite{Arkani-Hamed:2019ymq}
        N.~Arkani-Hamed, Y.~t.~Huang and D.~O'Connell,
        ``Kerr black holes as elementary particles,''
        JHEP \textbf{01}, 046 (2020)
        doi:10.1007/JHEP01(2020)046
        [arXiv:1906.10100 [hep-th]].
        224 citations counted in INSPIRE as of 13 Sep 2024
    \end{itemize}
\section*{Hard to be catagorized}
\begin{itemize}
    \item %\cite{Jakobsen:2021zvh}
    G.~U.~Jakobsen, G.~Mogull, J.~Plefka and J.~Steinhoff,
    ``SUSY in the sky with gravitons,''
    JHEP \textbf{01}, 027 (2022)
    doi:10.1007/JHEP01(2022)027
    [arXiv:2109.04465 [hep-th]].
    114 citations counted in INSPIRE as of 13 Sep 2024\\
    Presents a study of the gravitational scattering of two massive, spinning compact bodies, showing that this astrophysical system enjoys a hidden N = 2 supersymmetry.
    \item 
\end{itemize}
\end{document}