### 2025-03-19 Note/Note.tex
New

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx}

% 使用系统字体，需要 XeLaTeX 或 LuaLaTeX 编译
\usepackage{fontspec}
\setmainfont{Times New Roman}

\begin{document}

\title{Note for Scattering Amplitude Computation}
\author{Su Yingze}
\maketitle

\section{4-point case}
For the four point case, we can construt the color-ordered amplitude from the residue.

\end{document}