### 2025-05-09 Master Thesis/main.tex
New

\documentclass[11pt]{JHEP}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}

\title{Your Paper Title}
\author{Your Name\\
Your Institute\\
\email{your.email@domain}}

\abstract{
A short summary of your paper goes here. This includes background, results, and key findings.
}

\begin{document}

\section{Introduction}

This is the introduction to your JHEP-style article. You can cite as in~\cite{sample}.

\section{Main Results}

Equation example:
\begin{equation}
    E = mc^2
\end{equation}

\section{Conclusion}

Summarize what you found and its implications.

\bibliographystyle{JHEP}
\bibliography{references}

\end{document}